Information Aggregation in Exponential Family Markets

### 2014-02-12 WriteUp/maxentpredmkts.tex
returned the storkey reference
revised abstract and intro

\begin{abstract}
We consider the design of prediction market mechanisms known as automated market makers. We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a popular and well-studied probability distribution template used in statistics. We give a full development of this relationship and explore a range of benefits. We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that relying on exponential family distributions. We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders elicit risk aversion according to exponential utility. We also consider similar phenomena under alternative models, such as when traders are budget limited.
%
% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

@@

Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event. Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event. Traders are allowed to report their beliefs by buying and selling securities whose ultimate payoff depends on the future outcome. This will affect the state of the market, thus updating the predictive probabilities for the event. Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information. In some sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.

@@

We give an brief overview of the work on prediction market design below. Much of this work has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few. Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices. That is, how can we interpret the equilibrium market state when we have a number of traders with diverse beliefs on the underlying state of the world? In what sense is the market an aggregation mechanism? Do price changes relate to our usual Bayesian notion of information incorporation via posterior updating?

In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations. Common concepts in statistics -- e.g. \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference} -- relate to natural aspects of our class of mechanisms. In particular, the central objects in our market framework can be interpreted via conceptual objects used to define exponential families:
\begin{itemize}
	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
	\item the market prices correspond to \emph{mean parameters};
	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
\end{itemize}
% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}

We begin with a discussion of scoring rules in Section~\ref{sec:scoring} for exponential family distributions, and in particular we show that the logarithmic scoring rule is the only score that is \emph{proper} for the class of exponential families.
We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms. In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well. In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
\begin{itemize}
	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists. In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
	\item In Section~\ref{sec:exputil} we consider when our agents are risk-averse in an interesting special case, that is under the assumption they utilize \emph{exponential utility} to optimize their bets. We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. In this case we can characterize precisely how a single trader interacts with the market as well as the equilibrium reached given multiple traders; this result is achieved via using a \emph{potential game} argument. The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are set according to risk aversion parameters.
	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market. We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
\end{itemize}

@@

\textbf{Related Work.}
The notion of an exponential family distribution is fundamental to this paper. For a comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}. Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.

>>>>

\begin{abstract}
We consider the design of prediction market mechanisms known as automated market makers. We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a popular and well-studied probability distribution template used in statistics. We give a full development of this relationship and explore a range of benefits. We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that rely on exponential family distributions. We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility. We also consider similar aspects under alternative models, such as when traders are budget constrained.
%
% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

@@

Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event. Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event. Traders are allowed to report their beliefs by buying and selling securities whose ultimate payoff depends on the future outcome. This will affect the state of the market, thus updating the predictive probabilities for the event. Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information. In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.

@@

%We give an brief overview of the work on prediction market design below. 
Much of the work on prediction market design has focused heavily on structural properties of the mechanism: incentive compatibility, the market maker loss, the available liquidity, the fluctuations of the prices as a function of the trading volume, to name a few. Absent from much of the literature is a corresponding \emph{semantics} of the market behavior or the observed prices. That is, how can we interpret the equilibrium market state when we have a number of traders with diverse beliefs on the underlying state of the world? In what sense is the market an aggregation mechanism? Do price changes relate to our usual Bayesian notion of information incorporation via posterior updating?

In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations. Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}---relate to natural aspects of our class of mechanisms. In particular, the central objects in our market framework can be interpreted via concepts used to define exponential families:
\begin{itemize}
	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
	\item the market prices correspond to \emph{mean parameters};
	\item the market's cost function corresponds to the distribution's \emph{log-partition} function.
%	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
\end{itemize}
% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}

We begin in Section~\ref{sec:scoring} with a discussion of scoring rules based on exponential family distributions, and we show how the framework leads to a variety of scoring rules for continuous outcome spaces.
We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms. In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well. In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
\begin{itemize}
	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists. In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
	\item In Section~\ref{sec:exputil} we consider \emph{risk-averse} agents that optimize their bets according to exponential utility. 
	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; this result is achieved via a potential game argument. The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are proportional to risk aversion parameters.
	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market. We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
\end{itemize}

@@

\textbf{Related Work.}
The notion of an exponential family distribution is fundamental to this paper. For comprehensive introductions to these distributions, see~\citep{Barndorff78,WainJordan08}. Exponential families are intimately tied to the notions of log loss and entropy, but can be generalized to other types of convex losses and information, as shown by~\citet{GrunwaldDawid}, who also make a connection to scoring rules.
